---
title: "NYC Plant Data"
author: "charlee cobb"
date: "2/21/2023"
output: html_document
---
Read in csv of plant data and park code data
```{r}
plant_entitation <- read.csv("Entitation.csv")
parks <- read.csv("Parks_Properties.csv")

head(plant_entitation)
head(parks)
```


**Loking at only Forest Hills Area Parks**
From parks we need PROPNAME and GISPROPNUM of parks in Forest Hills area. 
Target parks: Flushing Meadows Corona Park, Forest Park, Maple Grove Cemetery, St.John Cemetery, Juniper Valley Park, Gardens Rodelberg Park, Hawthorne Park Forest Hills, Flag Pole Green, MacDonald Park, Yellowstone Park, Kissena Park, Cunningham Park
```{r}
library(dplyr)
target_parks <-c("Flushing Meadows Corona Park", "Forest Park", "Maple Grove Cemetery", "St.John Cemetery", "Juniper Valley Park", "Gardens Rodelberg Park", "Hawthorne Park Forest Hills", "Flag Pole Green", "MacDonald Park", "Yellowstone Park", "Kissena Park", "Cunningham Park")
l <- parks[parks$SIGNNAME %in% target_parks, ]
#subset the parks and only get property and gispropnum
forest_hills_area_parks <- select(l, GISPROPNUM, SIGNNAME)
forest_hills_area_parks %>% distinct()
```

Add a QUEID column based on the 'forest_hills_area_parks$GISPROPNUM' 
```{r}
#fix this, it's only pulling Q021
grep("Q015", plant_entitation$UEID)
ls <- forest_hills_area_parks$GISPROPNUM
ls

for(i in ls){
  if(grepl(i, plant_entitation$UEID, ignore.case = TRUE)){
    plant_entitation$QUEID <- i}
  else{
    print(i)
  }}



#checking to see if QUEID column and value are added
head(plant_entitation)
#check for forest park in QUEID
grep("Q015", plant_entitation$QUEID)

#merge dataframes, select QUEID, Parkname (SIGNNAME), 6 record names, 6 corresponding latin names, and 6 corresponding native status 
forest_hills_plant_entitation <- select(merge(plant_entitation, forest_hills_area_parks, by.x = "QUEID", by.y = "GISPROPNUM"),
                                        QUEID, SIGNNAME, RecordedName1, LatinName1, NativeStatus1, RecordedName2, LatinName2,
                                        NativeStatus2, RecordedName3, LatinName3, NativeStatus3, RecordedName4, LatinName4,
                                        NativeStatus4,
                                        RecordedName5, LatinName5, NativeStatus5, RecordedName6, LatinName6,
                                        NativeStatus6)


```


get only forest park out!
```{r}
library(MASS) 
library(reshape2) 
library(reshape)
library(dplyr)

#pull forest park data out of plant entitation. Doing this here because
#it didn't work in the above block, and we know "Q015" is Forest Park
forest_park <- plant_entitation[grep("Q015", plant_entitation$UEID), ]
head(forest_park)


#pull all record names, latin names, height, and native status
forest_park <- select(forest_park, RecordedName1, LatinName1, Height1, NativeStatus1, RecordedName2, LatinName2, Height2,
                                        NativeStatus2, RecordedName3, LatinName3, Height3, NativeStatus3, RecordedName4, LatinName4,
                                        Height4, NativeStatus4, RecordedName5, LatinName5, Height5, NativeStatus5, RecordedName6, LatinName6,
                                        Height6, NativeStatus6)
head(forest_park)


#make column names so we can melt the data into 4 columns: Common_Name", "Latin_Name", "Height", "Native_Status"
new_col_names <- c("Common_Name", "Latin_Name", "Height", "Native_Status",
                   "Common_Name", "Latin_Name", "Height", "Native_Status",
                   "Common_Name", "Latin_Name", "Height", "Native_Status",
                   "Common_Name", "Latin_Name", "Height", "Native_Status",
                   "Common_Name", "Latin_Name", "Height", "Native_Status",
                   "Common_Name", "Latin_Name", "Height", "Native_Status")
colnames(forest_park) <- new_col_names
head(forest_park)

#melt data
melt_forest_park <- melt(forest_park, id = c("Common_Name", "Latin_Name", "Height", "Native_Status"))
head(melt_forest_park)

#get rid of duplicates
forest_park_with_Height_plants_under_30 <- melt_forest_park[melt_forest_park$Height != ">30'", ]
no_repeats_forest_park_with_Height_plants_under_30 <- forest_park_with_Height_plants_under_30 %>% distinct()
write.csv(no_repeats_forest_park_with_Height_plants_under_30, "forest_park_plants_with_heights.csv")



#height enteries are really weired, there are multiple plant entries with different heights
#and height ranges, so just removing height all together
forest_park_no_height <- melt_forest_park[,c("Common_Name", "Latin_Name", "Native_Status")]
no_repeats_forest_park_no_height <- forest_park_no_height %>% distinct()
write.csv(no_repeats_forest_park_no_height, "forest_park_plants_without_heights.csv")

```
